\documentclass{sigchi}

% Arabic page numbers for submission.
% Remove this line to eliminate page numbers for the camera ready copy
%\pagenumbering{arabic}

% Load basic packages
\usepackage{balance}  % to better equalize the last page
\usepackage{graphics} % for EPS, load graphicx instead 
\usepackage[T1]{fontenc}
\usepackage{txfonts}
\usepackage{mathptmx}
\usepackage[pdftex]{hyperref}
\usepackage{color}
\usepackage{booktabs}
\usepackage{textcomp}
% Some optional stuff you might like/need.
\usepackage{microtype} % Improved Tracking and Kerning
% \usepackage[all]{hypcap}  % Fixes bug in hyperref caption linking
\usepackage{ccicons}  % Cite your images correctly!
\usepackage[utf8]{inputenc} % for a UTF8 editor only

\usepackage{tikz}
\usepackage{pgfplots}
\pgfplotsset{compat=1.12}

\usepackage[inline]{enumitem}

% Paper metadata (use plain text, for PDF inclusion and later
% re-using, if desired).  Use \emtpyauthor when submitting for review
% so you remain anonymous.
\def\plaintitle{An Exploration of Automated Grading of Complex Assignments}
\def\plainauthor{Chase Geigle, ChengXiang Zhai, Duncan Ferguson}
\def\emptyauthor{}
\def\plainkeywords{Automatic grading; ordinal regression; supervised
learning; learning to rank; active learning; text mining}

% llt: Define a global style for URLs, rather that the default one
\makeatletter
\def\url@leostyle{%
  \@ifundefined{selectfont}{
    \def\UrlFont{\sf}
  }{
    \def\UrlFont{\small\bf\ttfamily}
  }}
\makeatother
\urlstyle{leo}

% To make various LaTeX processors do the right thing with page size.
\def\pprw{8.5in}
\def\pprh{11in}
\special{papersize=\pprw,\pprh}
\setlength{\paperwidth}{\pprw}
\setlength{\paperheight}{\pprh}
\setlength{\pdfpagewidth}{\pprw}
\setlength{\pdfpageheight}{\pprh}

% Make sure hyperref comes last of your loaded packages, to give it a
% fighting chance of not being over-written, since its job is to
% redefine many LaTeX commands.
\definecolor{linkColor}{RGB}{6,125,233}
\hypersetup{%
  pdftitle={\plaintitle},
% Use \plainauthor for final version.
%  pdfauthor={\plainauthor},
  pdfauthor={\emptyauthor},
  pdfkeywords={\plainkeywords},
  bookmarksnumbered,
  pdfstartview={FitH},
  colorlinks,
  citecolor=black,
  filecolor=black,
  linkcolor=black,
  urlcolor=linkColor,
  breaklinks=true,
  hypertexnames=false
}

\begin{document}

\title{\plaintitle}

\numberofauthors{1}
\author{%
  \alignauthor{Chase Geigle, ChengXiang Zhai, Duncan Ferguson\\
    \affaddr{University of Illinois at Urbana-Champaign}\\
    $\{$geigle1,czhai,dcf$\}$@illinois.edu}
  }

\maketitle

\begin{abstract}
Automated grading is essential for scaling up learning. In this paper, we
conduct the first systematic study of how to automate grading of a complex
assignment using a medical case assessment as a test case. We propose to
solve this problem using a supervised learning approach and introduce three
general complementary types of feature representations of such complex
assignments for use in supervised learning.  We first show with empirical
experiments that it is feasible to automate grading of such assignments
provided that the instructor can grade a number of examples. We further
study how to integrate an automated grader with human grading and propose
to frame the problem as learning to rank assignments to exploit pairwise
preference judgments and use NDPM as a measure for evaluation of the
accuracy of ranking.  We then propose a sequential pairwise online active
learning strategy to minimize the effort of human grading and optimize the
collaboration of human graders and an automated grader.  Experiment results
show that this strategy is indeed effective and can substantially reduce
human effort as compared with randomly sampling assignments for manual
grading.
\end{abstract}

\keywords{\plainkeywords}

%\category{H.2.8}{Database Applications}{Data mining}

\input{sec-intro}

\input{sec-related}

\input{sec-assign}

\input{sec-method}

\input{sec-eval}

\input{sec-conclu}

\section{Acknowledgments}

This material is based upon work supported by the National Science
Foundation Graduate Research Fellowship Program under Grant Number
DGE-1144245.

\balance

\bibliographystyle{SIGCHI-Reference-Format}
\bibliography{auto-grading}
\end{document}
